---
title: "R Notebook"
output: html_notebook
---

## Working with the NHANES data

Throughout this chapter, you will use the NHANES dataset from the NHANES R package. These data are collected by the CDC and can be thought of as a random sample of US residents.

Before moving on to investigate particular variables, you'll have an opportunity to briefly explore the data in this exercise.

Instructions
 - Load the dplyr, ggplot2, and NHANES packages, one at a time.
 - Use the names() function to list the variables included in NHANES. (See ?NHANES for more info.)
 - Create a barplot for HomeOwn broken down by Gender. Use geom_bar(position = "fill") to compare relative frequencies.
 - Create a density plot for SleepHrsNight colored by SleepTrouble and faceted by HealthGen.
 - Note that because of the discrete nature of the data collection, the density should be smoothed a bit using adjust = 2.
 - Use facet_wrap(~ HealthGen) for the faceting, which will make it easier to visualize how the variables of interest vary by.
 
 
 
```{r}
# Load packages
library(dplyr)
library(ggplot2)
library(NHANES)
#install.packages("NHANES")

# What are the variables in the NHANES dataset?
names(NHANES)

# Create bar plot for Home Ownership by Gender
ggplot(NHANES, aes(x = Gender, fill = HomeOwn)) + 
  geom_bar(position = "fill") +
  ylab("Relative frequencies")

# Density for SleepHrsNight colored by SleepTrouble, faceted by HealthGen
ggplot(NHANES, aes(x =  SleepHrsNight, col = SleepTrouble)) + 
  geom_density(adjust=2) + 
  facet_wrap(~HealthGen)
```

Notice that the warning message for the densities and the grey portions of the bars both indicate a large number of missing observations in the dataset. If this were your dataset, it would be important to stop here and consider the cause of the missingness. For the course, we will now proceed without worrying about the missing observations.    

Randomly allocating samples

As seen in the video, you will now use R to randomly permute the observations and calculate a difference in proportions that could arise from a null distribution.

Using the NHANES dataset, let's investigate the relationship between gender and home ownership. Type ?NHANES in the console to get a description of its variables.

Instructions
The dplyr, ggplot2, and NHANES packages have been loaded for you.

Subset the NHANES dataset to consider only individuals whose home ownership status is either "Own" or "Rent". Save the result to homes.
Perform a single permutation to evaluate whether home ownership status (i.e. HomeOwn) differs between the "female" and "male" groups:
In your mutate() call, use sample() to shuffle home ownership status. Call this new variable HomeOwn_perm, a permuted version of HomeOwn.
Group by Gender.
In both the original data and in the permuted data, compute the proportion of individuals who own a home. Note that this will calculate proportions for both genders since you've grouped by the Gender variable in the line before it.
Using the diff() function, calculate the difference in proportion of home ownership for both prop_own_perm, the permuted data, and prop_own, the original data.

```{r}
# Subset the data: homes
homes <- NHANES %>%
  select(Gender, HomeOwn) %>%
  filter(HomeOwn %in% c("Own", "Rent"))

# Perform one permutation 
homes %>%
  mutate(HomeOwn_perm = sample(HomeOwn)) %>%
  group_by(Gender) %>%
  summarize(prop_own_perm = mean(HomeOwn_perm == "Own"), 
            prop_own = mean(HomeOwn == "Own")) %>%
  summarize(diff_perm = diff(prop_own_perm),
            diff_orig = diff(prop_own))
```

Randomization dotplot (n = 10)
100xp
As you saw in the video, natural variability can be modeled from shuffling observations around to remove any relationships that might exist in the population. This is done with the rep_sample_n() function from the oilabs package. Within it, you must specify arguments for the data (tbl), the sample size, the number of samples to take (reps), and whether sampling should be done with or without replacement (replace). The output includes a new column, replicate, which indicates the sample number. For example,

homes %>% 
  rep_sample_n(size = 5, reps = 3)
will return three samples of 5 observations from the homes dataset you created in the last exercise. The first 5 rows will have a value of 1 in the replicate column, the next 5 rows will have a value of 2, and so on. Note that the default value for the replace argument is FALSE.

In this exercise, you will permute the home ownership variable 10 times. By doing so, you will ensure that there is no relationship between home ownership and gender, so any difference in home ownership proportion for female versus male will be due only to natural variability.

Instructions
The dplyr, ggplot2, NHANES, and oilabs packages have been loaded for you and the homes dataset you created in the last exercise is available in your workspace.

Using rep_sample_n(), repeat the permutation process from the last exercise to get 10 null differences in proportions (i.e. reps = 10) using all observations in the homes dataset. Assign the result to homeown_perm.
Print the permuted differences in homeown_perm to the console.
Using geom_dotplot(), plot the differences in proportions obtained by shuffling the HomeOwn variable. Adjust the size of the dots by including binwidth = .001 in your call to geom_dotplot().

```{r, eval=FALSE, include=FALSE}
# Perform 10 permutations
homeown_perm <- homes %>%
  rep_sample_n(size = nrow(homes), reps = 10) %>%
  mutate(HomeOwn_perm = sample(HomeOwn)) %>%
  group_by(replicate,Gender) %>%
  summarize(prop_own_perm = mean(HomeOwn_perm == "Own"), 
            prop_own = mean(HomeOwn == "Own")) %>%
  summarize(diff_perm = diff(prop_own_perm),
            diff_orig = diff(prop_own)) # male - female

# Print differences to console
print(homeown_perm)

# Dotplot of 10 permuted differences in proportions
ggplot(homeown_perm, aes(x = diff_perm)) + 
  geom_dotplot(binwidth = 0.001)
```